The Analog Method for Precipitation Downscaling: Finding Better Analog Situations at a Sub-Daily Time Step

### 2016-11-16 Moving temporal window - HESS/template.tex
HESS paper under corrections.

Finding better analog situations does not obligatory imply better skills to predict the precipitation. Therefore, the impact of the MTW introduction, and thus the selection of other analog dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see Sect. \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC -- 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

@@

@@

Figure \ref{fig:changes_S1_analogs} presents the changes in the distributions of the S1 criterion for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analog situations for the Ulrichen station on the whole calibration period (1961-2008), due to introduction of the MTW. The precipitation target remains as before, that is centered on 18~h~UTC (6~h~UTC to 6~h~UTC the next day). The shapes of the distributions of the conventional approach and the MTW are similar, but the values of the analogy criteria are now reduced (shifted to the left), and therefore better. For analogs with higher ranks (e.g. $20^{th}$ or $40^{th}$), the difference between the two distributions is larger than for the first ranks, which means that the improvement increases with the rank of the analogs.

@@

The new performance scores are provided in Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI} for the 2Z and 2Z-2MI methods respectively. The gains, relatively to the static approach on the reduced archive, range from 0.57 to 2.14 points for the 2Z method and from 1.53 to 2.20 points for the 2Z-2MI method.

@@

@@

No relationship was found between the improvement of the score and the reduction of the S1 criteria, neither with the season.

@@

Both the analogy criteria (Sect. \ref{sec:influence_precip}) and the performance scores (Sect. \ref{sec:improvement_CRPSS_precip_threshold}) are improved to a greater extent for heavier precipitation events. This is likely due to the fact that higher precipitation events are a consequence of atmospheric conditions with a higher dynamism, such as weather disturbances, which have a well marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure center and the fronts change significantly during a day. We can therefore expect to improve more significantly these situations with a higher dynamism when introducing a MTW, as better matches to the target situation may be found. On the opposite, days with a low dynamism of the atmospheric circulation, such as anticyclonic situations, will not be radically improved by the introduction of the MTW.

@@

\subsection{Is a simple reconstructed precipitation archive useful for the MTW ?}

A proportional distribution is certainly the simplest approach that can be performed. It consists in allocating proportional parts of the original daily time series into a new moving average of 24~h-totals (Fig. \ref{fig:illustration_disaggregation}). When using this reconstructed time series on the shorter period, the method performance was degraded and was even below the conventional method without MTW (Table \ref{table:disaggregation_proportional} to compare to Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}). The benefit of a better selection of the analog situations is lost due to a precipitation archive of poor quality.

\subsection{Can a proxy variable help reconstruct a better precipitation archive ?}

The best proxy variable was identified by means of a correlation analyses (on non-zero values) with the 6-hourly precipitation time series on the period 1982-2007. Precipitable water and relative humidity (at 1000~hPa, 925~hPa or 850~hPa) were assessed at the four points surrounding the catchment, and by allowing a time lapse between both series in order to take into account the significant distance separating the weather stations and the reanalysis grid point (details not shown). The best proxy for the Zermatt station is the precipitable water at 45° N and 7.5° E, with a time offset of -6~h.

Table \ref{table:proxy_CRPSS} presents the CRPSS scores obtained by the reconstituted series using the proxy of precipitable water. A slight improvement can be obtained compared to the proportional distribution method, but it is still relatively small, and most of the benefit of the MTW is again lost. A more recent reanalysis archive with more accurate moisture variables might produce better proxies.

@@

The improvement of the circulation analogy was found to be more important for heavy precipitation events, which are generally related to more dynamic atmospheric situations. This dynamism implies more specific circulation patterns that are evolving more rapidly. Thus, a MTW is found to be of particular interest in this kind of situation, and benefits the prediction of extreme precipitation events.

@@

Perspectives of using the MTW rely on the availability of long precipitation series at a sub-daily time step and with high accuracy. First, these archives of high temporal resolution precipitation data are increasing now with time. The second source is the establishment of precipitation reanalysis at a country scale, or the use of reanalysis-driven RCMs (Regional Climate Models) or LAMs (Limited Area Models) on a long period. When such precipitation data will be available, the use of a MTW in the AM should be considered for any application, especially when the prediction quality of extreme events is important.

@@

>>>>

@@

Finding better analog situations does not necessarily imply better skills to predict the precipitation. Therefore, the impact of the MTW introduction, and thus the selection of other analog dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see Sect. \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC -- 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

\subsection{Reconstruction of a precipitation archive}
\label{sec:reconstruction}

A proportional distribution is certainly the simplest approach that can be performed. It consists in allocating proportional parts of the original daily time series into a new moving average of 24~h-totals (Fig. \ref{fig:illustration_disaggregation}).

@@

Figure \ref{fig:changes_S1_analogs} presents the changes in the distributions of the S1 criterion for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analog situations for the Ulrichen station on the whole calibration period (1961-2008), due to introduction of the MTW. The precipitation target remains as before, that is centered on 18~h~UTC (6~h~UTC to 6~h~UTC the next day). The shape of the distributions of the conventional approach and the MTW are similar, but the values of the analogy criteria are now reduced (shifted to the left), and therefore better. For analogs with higher ranks (e.g. $20^{th}$ or $40^{th}$), the difference between the two distributions is larger than for the first ranks, which means that the improvement increases with the rank of the analogs.

@@

The new performance scores (CRPSS) are provided in Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI} for the 2Z and 2Z-2MI methods respectively. The gains, relatively to the static approach on the reduced archive, range from 0.57 to 2.14 points for the 2Z method and from 1.53 to 2.20 points for the 2Z-2MI method.

@@

\subsection{Using a reconstructed archive}

When using this reconstructed time series on the shorter period, the method performance was degraded and was even below the conventional method without MTW (Table \ref{table:disaggregation_proportional} to compare to Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}).

The best proxy variable was identified by means of a correlation analyses (on non-zero values) with the 6-hourly precipitation time series on the period 1982-2007. Precipitable water and relative humidity (at 1000~hPa, 925~hPa or 850~hPa) were assessed at the four points surrounding the catchment, and by allowing a time lapse between both series in order to take into account the significant distance separating the weather stations and the reanalysis grid point (details not shown). The best proxy for the Zermatt station is the precipitable water at 45° N and 7.5° E, with a time offset of -6~h (Table \ref{table:proxy_correlations}).

Table \ref{table:proxy_CRPSS} presents the CRPSS scores obtained by the reconstituted series using the proxy of precipitable water.

@@

@@

Both the analogy criteria (Sect. \ref{sec:influence_precip}) and the performance scores (Sect. \ref{sec:improvement_CRPSS_precip_threshold}) are improved to a greater extent for heavier precipitation events. This is likely due to the fact that higher precipitation events are a consequence of atmospheric conditions with a higher dynamics, such as weather disturbances, which have a well marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure center and the fronts change significantly during a day. We can therefore expect to improve more significantly these situations with a higher dynamics when introducing a MTW, as better matches to the target situation may be found. On the opposite, days with a low dynamics of the atmospheric circulation, such as anticyclonic situations, will not be radically improved by the introduction of the MTW.

@@

\subsection{Relevance of the simple reconstructed precipitation archive}

The benefit of a better selection of the analog situations is lost due to a precipitation archive of poor quality.

A slight improvement can be obtained compared to the proportional distribution method, but it is still relatively small, and most of the benefit of the MTW is again lost. A more recent reanalysis archive with more accurate moisture variables might produce better proxies.

@@

The improvement of the circulation analogy was found to be more important for heavy precipitation events, which are generally related to more dynamic atmospheric situations. This dynamic implies more specific circulation patterns that are evolving more rapidly. Thus, a MTW is found to be of particular interest in this kind of situation, and benefits the prediction of extreme precipitation events.

@@

Perspectives of using the MTW rely on the availability of long precipitation series at a sub-daily time step and with high accuracy. First, these archives of high temporal resolution precipitation data are increasing now with time. The second source is the establishment of precipitation reanalysis at a country scale, or the use of reanalysis-driven RCMs (Regional Climate Models) or LAMs (Limited Area Models) on a long period. Even though outputs from these models might be biased or not accurate enough, information on the timing of the precipitation events could be useful in order to disaggregate stations timeseries. When precipitation data with accurate time distribution will be available, the use of a MTW in the AM should be considered for any application, especially when the prediction quality of extreme events is important.

@@

\begin{table}[htb]
	\caption{Value of the coefficient of determination between the reconstructed 6-hourly precipitation time series using the listed variables, and the accurate time series on the period 1982-2007. The grid points surrounding the region are the following: 1) 5° E, 47.5° N, 2) 5° E, 45° N, 3) 7.5° E, 47.5° N, 4) 7.5° E, 45° N. The highest coefficient of determination is indicated in bold.}
	\begin{center}
		\begin{tabular}{l c c c c c c}
			\hline
			\multirow{2}{*}{Variable} & \multirow{2}{*}{Point} &  \multicolumn{5}{c}{Time lapse} \\
			&  & -12 h & -6 h & 0 h & +6 h & +12 h \\ 
			\hline
			\multirow{ 4}{*}{RH1000} & 1 & 0.668 & 0.669 & 0.684 & 0.683 & 0.670 \\
			& 2 & 0.669 & 0.669 & 0.683 & 0.681 & 0.669 \\
			& 3 & 0.662 & 0.673 & 0.691 & 0.682 & 0.673 \\
			& 4 & 0.666 & 0.671 & 0.688 & 0.681 & 0.668 \\ \hline
			\multirow{ 4}{*}{RH925} & 1 & 0.672 & 0.673 & 0.684 & 0.684 & 0.675 \\
			& 2 & 0.674 & 0.674 & 0.683 & 0.682 & 0.672 \\
			& 3 & 0.662 & 0.673 & 0.691 & 0.682 & 0.673 \\
			& 4 & 0.666 & 0.671 & 0.689 & 0.681 & 0.668 \\ \hline
			\multirow{ 4}{*}{RH850} & 1 & 0.675 & 0.675 & 0.679 & 0.678 & 0.671 \\
			& 2 & 0.681 & 0.690 & 0.691 & 0.677 & 0.664 \\
			& 3 & 0.665 & 0.680 & 0.693 & 0.683 & 0.675 \\
			& 4 & 0.675 & 0.694 & 0.706 & 0.681 & 0.659 \\ \hline
			\multirow{ 4}{*}{TCW} & 1 & 0.688 & 0.687 & 0.667 & 0.655 & 0.652 \\
			& 2 & 0.697 & 0.699 & 0.669 & 0.644 & 0.644 \\
			& 3 & 0.686 & 0.708 & 0.689 & 0.655 & 0.648 \\
			& 4 & 0.696 & \textbf{0.721} & 0.696 & 0.643 & 0.636 \\ \hline
		\end{tabular}
	\end{center}
	\label{table:proxy_correlations}
\end{table}
